---
pdf_document: default
title: "Early exposure to language (in any modality) supports number concept development"

output:
  html_document: default
  pdf_document: default
---

#Load required packages
```{r echo=TRUE, results='hide'}
library(tidyverse) 
library(ggplot2)
library(colorspace)
library(stargazer)
library(MASS) # for polr function (ordinal logistic regression)
library(generalhoslem) # for testing model fit (lipsitz test and two others)
library(sure) # package for calculating residuals for ordinal logistic regression (https://journal.r-project.org/archive/2018/RJ-2018-004/RJ-2018-004.pdf)
library(sjmisc) # for easy recoding of variables
library(kableExtra)
library(lmtest)
library(broom)
library(beeswarm)
library(ggbeeswarm)
library(ggpubr)
library(table1)
library(VGAM)
library(forcats)
```

#Set working directory (this should be wherever your file is saved) & Import data file
```{r echo=FALSE, results='hide'}
#setwd("") 
setwd("/Users/emilycarrigan/Dropbox/Data Analysis Work/")
GiveN <- read.csv("GN_Smsummary_Lgall_KW_201106.csv", na.strings = c("N/A", "I/R", "", "TBD", "N/R"))
View(GiveN)
```

#Subset data
```{r results='hide'}
GiveN_subset <- subset(GiveN, GiveN$Including.in.Study.=='Yes' & GiveN$Including_GiveN.=='Yes' & GiveN$GN_All_Ceiling_Conservative!='' & GiveN$SES..8.66.!='') # some participant IDs in data file never tested, not tested on this task, or missing demographic data
View(GiveN_subset)
str(GiveN_subset) # check the data
summary(GiveN_subset) # check missing values for  variables
```

#Organize/create relevant variables
```{r}
#ORDER Language Groups the way I want them
GiveN_subset$LanguageGroup <- as.factor(factor(as.character(GiveN_subset$Group_4cat), levels = c("English Early", "ASL Early", "English Later", "ASL Later"), exclude=NA))

##Create Modality and Timing variables and add to dataframe
GiveN_subset$Language_Modality <- factor(ifelse(GiveN_subset$Group_4cat == "English Early" | GiveN_subset$Group_4cat == "English Later", "English", "ASL"), levels = c("English", "ASL"))

GiveN_subset$Language_Timing <- factor(as.character(GiveN_subset$Group_2cat), levels = c("Early", "Later"), exclude="") 

#CHECK whether any participants' "Language Tested" for a particular task does not match their Language_Modality group
GiveN_subset$Language_Match <- ifelse(as.character(GiveN_subset$Child_LanguageTested) == as.character(GiveN_subset$Language_Modality), "Match", "NOT MATCH")

#Convert Age of Language Exposure (currently in months) into years
GiveN_subset$AgeofLangExp_years <- GiveN_subset$Age.of.Exposure..mo./12

#Recode Race to have fewer categories
GiveN_subset$Race_recoded <- dplyr::recode(as.character(GiveN_subset$Race), 'Asian' = "Asian", 'Black or African American'="Black or African American", 'Mixed'="More than one", 'Other'="Other/Missing", 'Unsure, or prefer not to answer' = "Other/Missing", 'White'="White", .missing="Other/Missing")

#Recode Ethnicity to have fewer categories
GiveN_subset$Ethnicity_recoded <- dplyr::recode(as.character(GiveN_subset$Ethnicity), 'Hispanic or Latino' = "Hispanic or Latino", 'Not Hispanic or Latino'="Not Hispanic or Latino", 'Prefer not to answer'="Prefer not to answer/Missing", .missing="Prefer not to answer/Missing")


#CREATE BINARY Knower-level variable from ordinal measure in data file
GiveN_subset$GN_sm_2cat <- factor(ifelse(GiveN_subset$GN_Small_Ceiling._FinalSet == 6, "CP-knower", "NOT CP-knower"), levels = c("NOT CP-knower", "CP-knower"))

##CREATE PERCENT correct Give-N LARGE measure (second analysis) - drawn from data in 5 separate columns 
GiveN_subset <-  dplyr::mutate(GiveN_subset, t7_corr = ifelse(is.na(GiveN_subset$Target7_FinalSet_Correct.), GiveN_subset$Target7_1st_Response_Fish_Correct., GiveN_subset$Target7_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t9_corr = ifelse(is.na(GiveN_subset$Target9_FinalSet_Correct.), GiveN_subset$Target9_1st_Response_Fish_Correct., GiveN_subset$Target9_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t10_corr = ifelse(is.na(GiveN_subset$Target10_FinalSet_Correct.), GiveN_subset$Target10_1st_Response_Fish_Correct., GiveN_subset$Target10_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t12_corr = ifelse(is.na(GiveN_subset$Target12_FinalSet_Correct.), GiveN_subset$Target12_1st_Response_Fish_Correct., GiveN_subset$Target12_FinalSet_Correct.))
GiveN_subset <-  dplyr::mutate(GiveN_subset, t16_corr = ifelse(is.na(GiveN_subset$Target16_FinalSet_Correct.), GiveN_subset$Target16_1st_Response_Fish_Correct., GiveN_subset$Target16_FinalSet_Correct.))


#Calculate mean proportion correct for Give-N trials 7, 9, 10, 12, & 16 (only for kids who got all five trials)

# Code below will appropriately *not* calculate mean (puts "NA") if any trial from above columns has NA value--THIS IS CORRECT, bc doing a proportion correct out of different numbers of trials for different kids is a no-go
GiveN_subset <-  GiveN_subset %>% rowwise() %>% dplyr::mutate(GN_Lg_propcorr_NA=mean(c(t7_corr, t9_corr, t10_corr, t12_corr, t16_corr)))
```


#Summary Tables
```{r}
##Participant Demographics by Language Group
table1::label(GiveN_subset$Age) <- "Age (years)"
table1::label(GiveN_subset$Sex) <- "Sex"
table1::label(GiveN_subset$Race_recoded) <- "Race"
table1::label(GiveN_subset$Ethnicity_recoded) <- "Ethnicity"
table1::label(GiveN_subset$SES..8.66.) <- "SES"
table1::label(GiveN_subset$AgeofLangExp_years) <- "Age of Language Exposure (yrs)"

##Demographics (four groups)
table1::table1(~Age + Sex + Race_recoded + Ethnicity_recoded + SES..8.66. + AgeofLangExp_years | LanguageGroup, data = GiveN_subset, overall=F)

##Demographics by Timing
table1::table1(~Age + Sex + Race_recoded + Ethnicity_recoded + SES..8.66. + AgeofLangExp_years | Language_Timing, data = GiveN_subset, overall=F)

##Demographics by Modality
table1::table1(~Age + Sex + Race_recoded + Ethnicity_recoded + SES..8.66. + AgeofLangExp_years | Language_Modality, data = GiveN_subset, overall=F)

```


#Count/prop. of kids who are CP/NON-CP knowers, by Language Timing & Modality
```{r}
table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Timing)
prop.table(table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Timing), margin=2) #proportions calculated summing across columns

table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Modality)
prop.table(table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Modality), margin=2)#proportions calculated summing across columns
```



# Figure 2: BEESWARM BOXPLOT FOR GN-2-cat (CP/Non) w/dot colors by Language Timing & Language Modality
```{r}
jitter <- position_jitter(width=0, height=0.075)

#N in code below sums rows from table() functions output in previous block
a <- ggplot(data=GiveN_subset, mapping = aes(x=Language_Timing, y=Age)) + geom_boxplot() + geom_dotplot(mapping=aes(fill=GN_sm_2cat, color=GN_sm_2cat), method = "dotdensity", binaxis = "y", position = jitter, stackdir = "center", dotsize=0.5, stackratio = 1.25) + theme(text = element_text(size=11))  + xlab("Language Timing") + ylab("Age (Years)") +  scale_fill_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (50)", "CP-Knowers (126)")) + scale_color_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (50)", "CP-Knowers (126)"))


#N for code below is: summary() values - NA values from table()
b <- ggplot(data=GiveN_subset, mapping = aes(x=Language_Modality, y=Age)) + geom_boxplot() + geom_dotplot(mapping=aes(fill=GN_sm_2cat, color=GN_sm_2cat), binaxis = "y", position = jitter, stackdir = "center", dotsize=0.5, stackratio = 1.25) + theme(text = element_text(size=11))  + xlab("Language Modality") + ylab("Age (Years)") +    scale_fill_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (50)", "CP-Knowers (126)")) + scale_color_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (50)", "CP-Knowers (126)"))

ggarrange(a, b, ncol=2, nrow=1, common.legend = TRUE, labels="auto")

```




##EARLY-ONLY graphs & comparison (Chi-square)
```{r}
#Create dataframe with only early-exposed
GN_early <- subset(GiveN_subset, GiveN_subset$Language_Timing=="Early")

#table of CP-knower status vs group (modality)
table(GN_early$GN_sm_2cat, GN_early$Language_Modality)
prop.table(table(GN_early$GN_sm_2cat, GN_early$Language_Modality), margin=2) #proportions calculated summing across columns

#graph of CP-knower status by Age & Language Modality (Early-exposed only)
ggplot(data=GN_early, mapping = aes(x=Language_Modality, y=Age)) + geom_boxplot() + geom_dotplot(mapping=aes(fill=GN_sm_2cat, color=GN_sm_2cat), binaxis = "y", position = jitter, stackdir = "center", dotsize=0.5, stackratio = 1.25) + theme(text = element_text(size=11), legend.position = c(0.14, 0.87))  + xlab("Early Language Group") + ylab("Age (Years)") +    scale_fill_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (23)", "CP-Knowers (70)")) + scale_color_manual(name = "CP-Knower Status", values=c("tan2", "springgreen3"), labels=c("Non-CP-Knowers (23)", "CP-Knowers (70)"))

chisq.test(table(GN_early$GN_sm_2cat, GN_early$Language_Modality))
#CP-Knower achievement does NOT differ by Language Modality
```

```{r}
as.numeric(GiveN_subset$Language_Timing)
with(GiveN_subset, cor.test(SES..8.66., as.numeric(Language_Timing)))
with(GiveN_subset, wilcox.test(SES..8.66. ~ Language_Timing))

with(GiveN_subset, cor.test(Age, as.numeric(Language_Timing)))
with(GiveN_subset, wilcox.test(Age ~ Language_Timing))


as.numeric(GiveN_subset$Language_Modality)
with(GiveN_subset, cor.test(SES..8.66., as.numeric(Language_Modality)))
with(GiveN_subset, wilcox.test(SES..8.66. ~ Language_Modality))

with(GiveN_subset, cor.test(Age, as.numeric(Language_Modality)))
with(GiveN_subset, wilcox.test(Age ~ Language_Modality))
```


## GIVE-N QUANTITIES 1-6, 2-CATEGORY (CP-knowers and NOT CP-knowers) - LOGISTIC REGRESSION
```{r}
null <-glm(formula = GN_sm_2cat ~ 1, data = GiveN_subset, family = "binomial")
summary(null) #resid dev: 210.07   AIC 212.07

j <- glm(formula = GN_sm_2cat ~ SES..8.66. + Language_Modality + Age + Language_Timing, data = GiveN_subset, family = "binomial")
summary(j) #resid dev: 162.72   AIC 172.72
##Age & Timing sig., Modality & SES not.
exp(coef(j)) #ODDS RATIOS
	#Kids in the early exposed group are 2.23 times more likely to have achieved CP-knower status than kids in the later group

#MODEL FIT MEASURES
with(j, null.deviance - deviance)
with(j, df.null - df.residual)
with(j, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE))
#Model j fits sig. better than a null model, p<.001

#TESTING OTHER POSSIBLE MODELS
k <- glm(formula = GN_sm_2cat ~ SES..8.66. + Language_Modality + Age*Language_Timing, data = GiveN_subset, family = "binomial")
summary(k) #resid dev: 159.75   AIC 171.75  # Age only significant predictor
lmtest::lrtest(j, k) # model with Age * Timing interaction not significantly different from model without the interaction

h <- glm(formula = GN_sm_2cat ~ Language_Modality + Age + Language_Timing, data = GiveN_subset, family = "binomial") #model without SES
summary(h)  # residual deviance: 165.98  AIC: 173.98
lmtest::lrtest(j, h) # model without SES not significantly different from model with SES

w <- glm(formula = GN_sm_2cat ~ SES..8.66.+ Language_Modality + Age, data = GiveN_subset, family = "binomial") #model without Language Timing
summary(w) # residual deviance: 166.75  AIC: 174.75
lmtest::lrtest(j, w) # model without Language Timing IS significantly different from model with Language Timing (combined with AIC and residual deviance, model WITH Language Timing fits data better)



#MODEL DIAGNOSTICS for Model j
# from http://www.sthda.com/english/articles/36-classification-methods-essentials/148-logistic-regression-assumptions-and-diagnostics-in-r/ (this uses dplyr functions that are "retired"--gather()--so I've modified below)

#Step 1: Checking that continuous predictors ARE in fact linearly associated with logit of the outcome

# Select only numeric predictors
continuous_pred <- GiveN_subset %>%
  dplyr::select(c("Age", "SES..8.66.")) 
predictors <- colnames(continuous_pred)

#get the predicted probabilities for EACH KID in the whole model (based on their specific values of each predictor)
probabilities <- data.frame(predict(j, type = "response"))

##**1/28/21 NEED TO FIX COLNAME AND ERROR: "Column `logit` must be length 1 (the group size), not 176"**

# Bind the logit and tidying the data for plot
continuous_pred <- continuous_pred %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  pivot_longer(names_to = "predictors", values_to = "predictor.value", -logit)

#scatterplots!
ggplot(continuous_pred, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") + 
  theme_bw() + 
  facet_wrap(~predictors, scales = "free_y")



# Step 2: Checking for "INFLUENTIAL VALUES" - extreme data points
plot(j, which = 4, id.n = 3)

# Extract model results
j.data <- augment(j) %>% 
  mutate(index = 1:n()) 

j.data %>% top_n(3, .cooksd)

ggplot(j.data, aes(index, .std.resid)) + 
  geom_point(aes(color = GN_sm_2cat), alpha = .5) +
  theme_bw()

j.data %>% 
  filter(abs(.std.resid) > 3)
#NO INFLUENTIAL OBSERVATIONS IN DATA


##Step 3: Check for Multicollinearity
car::vif(j)  ##None! (bc values not >5 or 10)

# WE HAVE A GOOD MODEL, FOLKS!

#Output Model J
stargazer(j, title = "Logistic Regression Results", align=TRUE, dep.var.labels=c("CP-Knower Status"), covariate.labels=c("Socioeconomic Status (SES)", "Language Modality (ASL)", "Age (Years)", "Timing of Language Exposure (Early)"), single.row=TRUE, keep.stat=c("n", "aic"),  out= "GN_CPknowerbinary_SES_Modality_Age_Timing_200514.htm")
```           


















0


##GIVE-N QUANTITIES 7, 9, 10, 12, 16

Checking number of children who were not CP-knowers but were tested on "large" quantities
```{r}
notCPknowers <- subset(GiveN_subset, GiveN_subset$GN_sm_2cat=="NOT CP-knower")
nonCP_GNlg <- subset(notCPknowers, !is.na(notCPknowers$GN_Lg_propcorr_NA)) #19 particpants
```

## FIGURE 3: SCATTERPLOTS of proportion correct values on GIVE-N Quantities 7, 9, 10, 12, 16 by Age (years), separated by Language Timing & Language Modality
```{r}
GN_lg_noNA <- subset(GiveN_subset, GiveN_subset$GN_Lg_propcorr_NA!='' & GiveN_subset$GN_sm_2cat!="NOT CP-knower")
summary(GN_lg_noNA$Language_Timing)
table(GN_lg_noNA$Language_Timing, GN_lg_noNA$GN_Lg_propcorr_NA, useNA = "always")

#N for code below is: summary() values - NA values from table()
GNlg_Timing <- ggplot(data=GN_lg_noNA, aes(x=Age, y=GN_Lg_propcorr_NA, shape=Language_Timing, color=Language_Timing)) + geom_jitter(na.rm=TRUE) + geom_smooth(method=lm, se = FALSE) + labs(x="Age at Test (Years)", y="Give-N Large Proportion Correct") + theme(plot.title = element_text(hjust = 0.5), legend.position = c(0.82, 0.15)) + scale_shape_manual(name="Language\nTiming", values=c(17, 16), labels=c("Early (68)", "Later (48)")) + scale_color_manual(name="Language\nTiming", values=c("deepskyblue3", "firebrick3"), labels=c("Early (68)", "Later (48)"))


summary(GN_lg_noNA$Language_Modality)
table(GN_lg_noNA$Language_Modality, GN_lg_noNA$GN_Lg_propcorr_NA, useNA = "always")

#N for code below is: summary() values - NA values from table()
GNlg_Modality <- ggplot(data=GN_lg_noNA, aes(x=Age, y=GN_Lg_propcorr_NA, shape=Language_Modality, color=Language_Modality)) + geom_jitter(na.rm=TRUE) + geom_smooth(method=lm, se = FALSE) + labs(x="Age at Test (Years)", y="Give-N Large Proportion Correct") + theme(plot.title = element_text(hjust = 0.5), legend.position = c(0.81, 0.15)) + scale_color_manual(name = "Language\nModality", values=c("slategray4", "purple3"), labels=c("English (73)", "ASL (43)")) + scale_shape_manual(name = "Language\nModality", values=c(19, 15),  labels=c("English (73)", "ASL (43)"))

ggarrange(GNlg_Timing, GNlg_Modality, ncol=2, nrow=1, common.legend = FALSE, labels="auto")

```

Both linear and tobit models had poor measures of model fit and poor residual plots, so using an ordinal model
Model *does not* include the 19 participants who were tested on large quantities despite not being CP-knowers
```{r}
#ORDINAL MODEL

GN_lg_noNA$GN_lg_factor <- as.factor(as.character(GN_lg_noNA$GN_Lg_propcorr_NA))

GN_lg_null <- polr(formula = GN_lg_factor ~ 1, data = GN_lg_noNA,  method = "probit", Hess = TRUE) 
summary(GN_lg_null) #res dev. 305  AIC: 315

GN_lg_polr <- polr(formula = GN_lg_factor ~ SES..8.66. + Language_Modality + Age + Language_Timing, data = GN_lg_noNA,  method = "probit", Hess = TRUE) 
summary(GN_lg_polr)  # res dev 265   AIC: 283
tidy(GN_lg_polr, conf.int = TRUE, p.value=TRUE)
dropterm(GN_lg_polr)

lipsitz.test(GN_lg_polr) #ns
logitgof(GN_lg_noNA$GN_lg_factor, fitted(GN_lg_polr)) #ns
pulkrob.chisq(GN_lg_polr, c("Language_Modality", "Language_Timing")) #ns

## checking RESIDUALS

sres <- resids(GN_lg_polr)

p1 <- autoplot(sres, what = "qq", distribution = qnorm)
p2 <- autoplot(sres, what = "covariate", x = GN_lg_noNA$SES..8.66., xlab = "SES") #bit of bunching at higher values but we know SES is skewed (also not variable of interest in this model)
p3 <- autoplot(sres, what = "covariate", x = GN_lg_noNA$Language_Modality, xlab = "Language Modality") #good
p4 <- autoplot(sres, what = "covariate", x = GN_lg_noNA$Age, xlab = "Age") #good
p5 <- autoplot(sres, what = "covariate", x = GN_lg_noNA$Language_Timing, xlab = "Language Timing") #good
p6 <- autoplot(sres, what = "covariate", x = GN_lg_noNA$GN_lg_factor, xlab = "Large Number Knowledge") # shows that surrogate residuals somewhat underestimated for lower values of Give-N large prop corr (0, 0.2, 0.4, 0.6, 0.8) and overestimated for highest values (likely indicating that ordinal characterization of outcome not great, but this is not the main point of our paper)
grid.arrange(p1, p2,p3, p4, p5, p6, ncol = 3)  # code that produces Figure 2 in Greenwell et al 2019 (R Journal)

## Decent fitting model


stargazer(GN_lg_polr, title = "Ordinal Logistic Regression Results", align=TRUE, dep.var.labels=c("Give-N Large Proportion Correct"), covariate.labels=c("Socioeconomic Status (SES)", "Language Modality (ASL)", "Age (Years)", "Timing of Language Exposure (Later)"), single.row=TRUE, keep.stat=c("n", "aic"),  out= "GN_lgordinal_SES_Modality_Age_Timing_probit_201228.htm")
```

Number of kids in GN small vs. GN large models (divided by Group, Language Timing, and Language Modality)
```{r}
addmargins(table(GiveN_subset$GN_sm_2cat, GiveN_subset$LanguageGroup))
addmargins(table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Timing))
addmargins(table(GiveN_subset$GN_sm_2cat, GiveN_subset$Language_Modality))


addmargins(table(GN_lg_noNA_2$GN_Lg_propcorr_NA, GN_lg_noNA_2$LanguageGroup))
addmargins(table(GN_lg_noNA_2$GN_Lg_propcorr_NA, GN_lg_noNA_2$Language_Timing))
addmargins(table(GN_lg_noNA_2$GN_Lg_propcorr_NA, GN_lg_noNA_2$Language_Modality))

```